---
title: "Getting started with cluster analysis using *transcluster*"
author: "James Stimson"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

*transcluster* implements new methods for the partition of data obtained from whole genome sequencing (WGS) studies into transmission clusters. Inputs include:

- Sample unique IDs
- Sample dates
- Matrix of pairwise SNP differences between the samples
- Resistance conferring SNP data

## Functionality

The available functions in *transcluster* are:

**withinCluster**: test whether two cases are within the same transmission cluster

**createModel**: create model to be used for cluster analysis

**makeSNPClusters**: make SNP-based clusters for each threshold set in the model

**makeTransClusters**: make transmission-based clusters for each threshold set in the model

**setDatesFromFile**: set sample dates from a file

**setParams**: set parameters for existing cluster analysis model

**setSNPFromFile**: set SNP distance matrix from a file

**setSNPThresholds**: set SNP thresholds for existing cluster analysis model

**setTransMatrixFromFile**: set transmission distance matrix from a file

**setTransThresholds**: set transmission thresholds for existing cluster analysis model

## A simple example

Start by loading the libraries:
```{r}
library(transcluster)
library(stats)
library(clue)
library(igraph)
```

Then create some data. This is the same as the first example in the paper.
```{r}
ids <- c('A', 'B', 'C', 'D')
dates <- c(2018, 2014, 2016, 2016)
snpMatrix <- matrix(c(0,5,7,7,5,0,8,8,7,8,0,6,7,8,6,0), nrow=4, ncol=4)
```

Set up the model:
```{r}
myModel <- createModel(ids, dates, snpMatrix)
myModel <- setParams(myModel, lambda=1.5, beta=2.3)
myModel <- setSNPThresholds(myModel, c(5,6,7))
myModel <- setTransThresholds(myModel, c(7,8,9))
```

Create the clusters:
```{r}
mySNPClusters <- makeSNPClusters(myModel, 'test')
myTransClusters <- makeTransClusters(myModel, 'test')
```

Examine the results using igraph:
```{r cache=FALSE }
library(transcluster)
myModel <- setCutoffs(myModel)
plotClusters(myTransClusters[['7']], myModel, vSize=40, vFontSize=1)
plotClusters(myTransClusters[['8']], myModel, vSize=40, vFontSize=1)
plotClusters(myTransClusters[['9']], myModel, vSize=40, vFontSize=1)
```

And compare these to the clusters obtained with SNP-based clustering:
```{r}
plotClusters(mySNPClusters[['5']], myModel, vColor='orange', vSize=40, vFontSize=1)
plotClusters(mySNPClusters[['6']], myModel, vColor='orange', vSize=40, vFontSize=1)
plotClusters(mySNPClusters[['7']], myModel, vColor='orange', vSize=40, vFontSize=1)
```

The results also appear in **.csv** files saved locally.

## More examples

Test directly whether two cases are in the same cluster:
```{r}
withinCluster(N = 3, k_cutoff = 5, t1 = 2012.13, t2 = 2015.76, lambda = 0.7, beta = 1.3, perc_cutoff = 0.2)
```


## Notes

More interesting examples to come^[Under construction.].


